---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data  
The first step is to unzip the given file "activity.zip"
```{r first_step}
unzip("activity.zip")
```

Then it is read into R and its contents examined
```{r reading_into_R}
dat <- read.csv("activity.csv")
head(dat)
tail(dat)
dim(dat)
```

**Data Processing**  
Below is the steps taken to prepare the data for the assignment
```{r processing}
library(lubridate)
library(tidyverse)
dat$date <- ymd(dat$date)
dat1 <- dat %>%
        filter(!is.na(steps)) %>%
        group_by(date) %>%
        mutate(total_steps = sum(steps))
histo <- data.frame(date=unique(dat1$date), total_steps=unique(dat1$total_steps))
```

**Plotting**  
The following histogram shows the frequency distribution of the daily total steps
```{r plotting}
ggplot(data = histo) +
        geom_histogram(aes(total_steps), binwidth = 600,
                       color = "black", fill = "sky blue") +
        xlab("Daily Total Steps") +
        ylab("Frequency")
```



## What is mean total number of steps taken per day?  
**Calculating "mean" and "median"**  
The mean and median of the total numbers of steps taken per days are calculated below
```{r mean and median}
cat("Mean is", mean(histo$total_steps))
cat("Median is", median(histo$total_steps))
```



## What is the average daily activity pattern?
**Time Series Plot**  
```{r Time_Series_Plot}
daily <- dat1 %>%
        ungroup %>%
        group_by(interval) %>%
        mutate(interval_mean = mean(steps)) %>%
        arrange(desc(interval_mean))
ggplot(daily, aes(interval, interval_mean)) +
        geom_line(color = "Purple") +
        labs(x = "Interval", y = "Average Daily Steps") +
        ggtitle("Time Series Plot")
```

**The 5-minute interval which contains the maximum number of steps is**
```{r max_interval_mean}
cat("The 5-minute interval with maximum steps is", daily$interval[1])
```


## Imputing missing values
**Imputation of missing values are done by mean of th interval across all days**
```{r imputation}
cat("The total number of missing value is,", sum(is.na(dat)))
dat2 <- dat %>%
        group_by(interval) %>%
        mutate(interval_mean = mean(steps, na.rm = TRUE))
imputed <- data.frame(sapply(dat2,
                             function(x) ifelse(is.na(x),
                                                dat2$interval_mean, x)))
imputed$steps <- as.double(imputed$steps)
imputed$date <- as.POSIXlt.Date(imputed$date)
imputed$interval_mean <- as.double(imputed$interval_mean)

head(imputed)
tail(imputed)
dim(imputed)
```

**Histogram from Imputed Data frame**
```{r histogram}
imputed$date <- ymd(imputed$date)

dat3 <- imputed %>%
  group_by(date) %>%
  mutate(total_steps = sum(steps))

histo_imputed <- dat3[!duplicated(dat3$date),]

ggplot(data = histo_imputed) +
        geom_histogram(aes(total_steps), binwidth = 600,
                       color = "black", fill = "sky blue") +
        xlab("Daily Total Steps") +
        ylab("Frequency")

cat("Mean after imputation is", mean(dat3$total_steps))
cat("Median after imputation is", median(dat3$total_steps))
```

After imputation, median and mean became equal.  




## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays}
week_days <- dat3 %>%
  ungroup() %>%
  mutate(is_weekend = ifelse((weekdays(date)=="Saturday" | weekdays(date)=="Sunday"),
                             "Weekend", "Weekdays"))
week_days$is_weekend <- as.factor(week_days$is_weekend)
dat4 <- week_days %>%
  group_by(interval, is_weekend) %>%
  mutate(interval_mean = mean(steps))

ggplot(dat4, aes(interval, interval_mean, color = is_weekend)) +
  geom_line() +
        labs(x = "Interval", y = "Average Daily Steps") +
        ggtitle("Time Series Plot after Imputation") +
  facet_wrap(. ~ is_weekend)
```

