---
title: "Activity_Monitoring"
author: "Naing Aung"
date: "8/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Course Assignment Submission

The first step is to unzip the given file "activity.zip"
```{r first step}
unzip("activity.zip")
```

Then it is read into R and its contents examined
```{r reading into R}
dat <- read.csv("activity.csv")
head(dat)
tail(dat)
dim(dat)
```

**Data Processing**  
Below is the steps taken to prepare the data for the assignment
```{r processing}
library(lubridate)
library(tidyverse)
dat$date <- ymd(dat$date)
dat$interval <- as.factor(dat$interval)
dat <- dat %>%
        filter(!is.na(steps)) %>%
        group_by(date) %>%
        mutate(total_steps = sum(steps),
               average_daily_steps = mean(steps))
dat1 <- data.frame(date=unique(dat$date), total_steps=unique(dat$total_steps))
```

**Plotting**  
The following histogram shows the frequency distribution of the daily total steps
```{r plotting}
ggplot(data = dat1) +
        geom_histogram(aes(total_steps), binwidth = 600,
                       color = "black", fill = "sky blue") +
        xlab("Daily Total Steps") +
        ylab("Frequency")
```

**Calculating "mean" and "median"**  
The mean and median of the total numbers of steps taken per days are calculated below
```{r mean and median}
cat("Mean is", mean(dat1$total_steps))
cat("Median is", median(dat1$total_steps))
```

**Time Series Plot**  
```{r Time Series Plot}

```


